---
title: "MNIST Exploratory Analysis"
output: html_document
---

[Return to main page](index.html)

```{r setup, echo=FALSE, message=FALSE}
library(plotly) # Interactive data visualization
library(ggplot2) # Data visualization
library(Rtsne) # t-SNE
```

### Exploratory Data Analysis {.tabset}

```{r MNIST Data Preparation, echo = FALSE, eval = FALSE}
# DATA INGESTION
mnistData <- dslabs::read_mnist()

# DATA FORMATTING
mnistData$train$labels <- forcats::as_factor(mnistData$train$labels)
mnistData$test$labels <- forcats::as_factor(mnistData$train$labels)

saveRDS(mnistData, "../assets/Data-Objects/mnistData.rds")
```


#### Principal Components Analysis

```{r MNIST PCA, echo = FALSE}
mnistData <- readRDS("../assets/Data-Objects/mnistData.rds")
mnistPCA <- princomp(mnistData$train$images, cor = FALSE)
mnistPCA$scores <- as.data.frame(mnistPCA$scores)
PVE <- mnistPCA$sdev^2 / sum(mnistPCA$sdev^2) * 100
```

```{r MNIST PCA Biplot, echo = FALSE, fig.align = "center", fig.width = 9}
ggplot(mnistPCA$scores) +
	geom_point(aes(Comp.1, Comp.2, color = mnistData$train$labels)) +
	labs(title = "Biplot",
		 x = "Principal Component 1 (PVE = 9.7%)",
		 y = "Principal Component 2 (PVE = 7.1%)",
		 color = "Digit") +
	theme_bw() +
ggplot(mnistPCA$scores) +
	geom_density_2d(aes(Comp.1, Comp.2, color = mnistData$train$labels)) +
	labs(title = "Density Biplot",
		 x = "Principal Component 1 (PVE = 9.7%)",
		 y = "Principal Component 2 (PVE = 7.1%)",
		 color = "Digit") +
	theme_bw() +
patchwork::plot_layout(axes = "collect")
```

```{r MNIST PCA 3D Plot, echo = FALSE, fig.align = "center", fig.width = 9, warning = FALSE}
plotly::plot_ly(mnistPCA$scores, x = ~Comp.1, y = ~Comp.2, z = ~Comp.3,
	type = "scatter3d", mode = "markers", color = mnistData$train$labels) |>
	plotly::layout(title = "\nPCA 3D Plot",
		scene = list(xaxis = list(title = "PC 1 (PVE = 9.7%)"),
		yaxis = list(title = "PC 2 (PVE = 7.1%)"),
		zaxis = list(title = "PC 3 (PVE = 6.2%)")))
```

It is evident that linear projections only capture some of the structure of the
images and that we need to extend our dimensionality reduction efforts. To
accomplish this, we will leverage the $t$-distributed Stochastic Neighbor
Embedding ($t$-SNE) algorithm.

#### $t$-distributed Stochastic Neighbor Embedding


```{r MNIST t-SNE, echo = FALSE, eval = FALSE}
mnist2DTSNE <- Rtsne::Rtsne(mnistData$train$images, dims = 2, num_threads = 0)
saveRDS(mnist2DTSNE, "../assets/Data-Objects/tsne_2D_perplexity_30.rds")

mnist3DTSNE <- Rtsne::Rtsne(mnistData$train$images, dims = 3, num_threads = 0)
saveRDS(mnist3DTSNE, "../assets/Data-Objects/tsne_3D_perplexity_30.rds")
```

```{r MNIST t-SNE Biplots, echo = FALSE, fig.align = "center", fig.width = 9}
mnistTSNE <- readRDS("../assets/Data-Objects//tsne_2D_perplexity_30.rds")
ggplot(as.data.frame(mnistTSNE$Y)) +
	geom_point(aes(V1, V2, color = mnistData$train$labels)) +
	labs(title = "MNIST t-SNE Biplot",
		 x = 'x',
		 y = 'y',
		 color = "Digit") +
	theme_bw() +
ggplot(as.data.frame(mnistTSNE$Y)) +
	geom_density_2d(aes(V1, V2, color = mnistData$train$labels)) +
	labs(title = "MNIST t-SNE Biplot",
		 x = 'x',
		 y = 'y',
		 color = "Digit") +
	theme_bw() +
patchwork::plot_layout(axes = "collect")
```

```{r MNIST t-SNE 3D Plot, echo = FALSE, fig.align = "center", fig.width = 9, warning = FALSE}
mnistTSNE <- readRDS("../assets/Data-Objects//tsne_3D_perplexity_30.rds")
plotly::plot_ly(as.data.frame(mnistTSNE$Y), x = ~V1, y = ~V2, z = ~V3,
	type = "scatter3d", mode = "markers", color = mnistData$train$labels) |>
	plotly::layout(title = "\nt-SNE 3D Plot",
	scene = list(xaxis = list(title = "Embedding Dimension 1"),
	yaxis = list(title = "Embedding Dimension 2"),
	zaxis = list(title = "Embedding Dimension 3")))
```
